Monte Carlo and bootstrap for ALL parameters with tables

### 2024-09-12 Intermediary_steps/adapt_boot2.Rmd
Better organized

```{r}
# Number of trees
random_points$var <- mapply(circle3_nb, random_points[, 1], random_points[, 2], MoreArgs = list(15, 9, 6))
est_init_nb <- sum(random_points$var/p)
cat("Number of trees:", nrow(trees), "\nEstimation of the number of trees:", est_init_nb)

@@

```{r}
# We only keep x and y in the dataframe
random_points_1 <- random_points
random_points_1$l <- NULL
```

@@

# Comparison between both methods

```{r}
comparison2 <- function(df_init, B, small_step) {
  # df_init: initial data frame with 4 variables, just as random_points
  # B: number of bootstrap replications each time
  # small_step: step to decide how many bootstrap series will be computed
  # large small_step => a few bootstrap series
  mean_vec1 <- c()
  mean_vec2 <- c()
  var_vec1 <- c()
  var_vec2 <- c()
  b1 <- 100
  while (b1  < B) {
    ## First method
    # We only keep x and y in the dataframe
    df_init_1 <- df_init
    df_init_1$l <- NULL
    df_boot <- f_boot(df_init_1, b1)
    est_boot_1 <- est_boot1(df_boot)

@@

    # tot_mean <- append(tot_mean, as.numeric(abs(mean(est_boot_2)-30942) <  abs(mean(est_boot_1)-30942))*2-1)
    # 1: the second method is better
    # tot_var <- append(tot_var, as.numeric(var(est_boot_2) <  var(est_boot_1))*2-1)
    # 1: the second method is better
    b1 <- b1+small_step
  }
  return(list(mean_vec1, mean_vec2, var_vec1, var_vec2))
}
```

@@

```{r}
df_test <- random_points
res2 <- comparison2(df_test, B_test, 1)
```

@@

p <- ggplot(data, aes(data[,1], data[,2] - data[,3]))
p + geom_point() +ggtitle("Comparison of the mean of estimators") + xlab("Number of bootstrap replications") + ylab("First method minus Second method")
```

```{r}
data_var = as.data.frame(cbind(1:length(res2[[1]]), res2[[3]], res2[[4]]))

p <- ggplot(data_var, aes(data_var[,1], data_var[,2] - data_var[,3]))
p + geom_point() + ggtitle("Comparison of the variance of estimators") + xlab("Number of bootstrap replications") + ylab("First method minus Second method")

# If it is positive, method 2 is better
```

```{r}
# To what extent does the second method improve variance?
cat("If it is positive, it indicates that the second method improve reducing the variance.")
cat("\nWith all bootstrap samples:", mean(data_var[,2] - data_var[,3]))
cat("\nWith the first half of bootstrap series (between 100 and ", floor((B_test+100)/2)," samples): ", mean(data_var[1:(floor((B_test+100)/2)-100),2] - data_var[1:(floor((B_test+100)/2)-100),3]), sep = "")
cat("\nWith the last half of bootstrap series (between ", floor((B_test+100)/2)," and ", B_test, " samples): ", mean(data_var[(floor((B_test+100)/2)-100):(B_test-100),2] - data_var[(floor((B_test+100)/2)-100):(B_test-100),3]), sep = "")

>>>>

```{r}
# Number of trees
# Local variable for each sampled point
random_points$var <- mapply(circle3_nb, random_points[, 1], random_points[, 2], MoreArgs = list(15, 9, 6))
est_init_nb <- sum(random_points$var/p)
cat("Number of trees:", nrow(trees), "\nEstimation of the number of trees:", est_init_nb)

@@

```{r}
# We only keep x, y and var in the dataframe
random_points_1 <- random_points
random_points_1$l <- NULL
```

@@

# Comparison between the two methods

```{r}
comparison2 <- function(df_init, B, small_step) {
  # df_init: initial data frame with 4 variables, just as random_points
  # B: maximum number of bootstrap replications for the last series
  # small_step: step to decide how many bootstrap series will be computed
  # large small_step => a few bootstrap series
  mean_vec1 <- c()
  mean_vec2 <- c()
  var_vec1 <- c()
  var_vec2 <- c()
  b1 <- 100 # minimum number of bootstrap samples for one single series
  while (b1  < B) {
    ## First method
    # We only keep x, y and var in the dataframe
    df_init_1 <- df_init
    df_init_1$l <- NULL
    df_boot <- f_boot(df_init_1, b1)
    est_boot_1 <- est_boot1(df_boot)

@@

    b1 <- b1+small_step
  }
  return(list(mean_vec1, mean_vec2, var_vec1, var_vec2))
}
```

@@

```{r}
df_test <- random_points

start.time <- Sys.time()

res2 <- comparison2(df_test, B_test, 1)

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

@@

p <- ggplot(data, aes(data[,1]+100, data[,2] - data[,3]))
p + geom_point() +ggtitle("Comparison of the mean of estimators") + xlab("Number of bootstrap replications") + ylab("First method minus Second method")
```
Figure 2. **Comparison of estimators obtained through the two methods depending on the number of bootstrap replications**. What is plotted is the gap between the result following the first method and the true value *minus* the gap between the result following the second method and the true value, as a function of the number of bootstrap replications. If it is positive, the second method gives estimators that are closer to the true value.

```{r}
# We look at the variance of the estimators
data_var = as.data.frame(cbind(1:length(res2[[1]]), res2[[3]], res2[[4]]))

p <- ggplot(data_var, aes(data_var[,1]+100, data_var[,2] - data_var[,3]))
p + geom_point() + ggtitle("Comparison of the variance of estimators") + xlab("Number of bootstrap replications") + ylab("First method minus Second method")

# If it is positive, method 2 is better
```
Figure 3. **Comparison of the variance of the estimators obtained through the two methods depending on the number of bootstrap replications**. What is plotted is the variance obtained by following the first method *minus* the variance obtained by following the second method, as a function of the number of bootstrap replications. If it is positive, the second method gives estimators that have smaller variances.

```{r}
# To what extent does the second method improve variance?
cat("If it is positive, it indicates that the second method helps reducing the variance.")
cat("\nWith all bootstrap samples:", mean(data_var[,2] - data_var[,3]))
cat("\nWith the first half of bootstrap series (between 100 and ", floor((B_test+100)/2)," samples): ", mean(data_var[1:(floor((B_test+100)/2)-100),2] - data_var[1:(floor((B_test+100)/2)-100),3]), sep = "")
cat("\nWith the last half of bootstrap series (between ", floor((B_test+100)/2)," and ", B_test, " samples): ", mean(data_var[(floor((B_test+100)/2)-100):(B_test-100),2] - data_var[(floor((B_test+100)/2)-100):(B_test-100),3]), sep = "")

### 2024-09-12 Intermediary_steps/test_boot_short_2.Rmd
Better organized

---
title: "Testing some bootstrap process"
output: pdf_document
date: "2024-08-02"
---

@@

Idea for the bootstrap: for each line l, containing $n_l$ sampled points, $n_l - 1$ points are sampled with replacement. As a result, $100(n_l-1)$ points are sampled from the initial sample. In this case, 900 points are sampled.

@@

for (l in 0:((1000/c_dim)-1)) {
  if (l==0){
    x_init <- i*c_dim
  }else if ((3*l+i)*c_dim < (10)*c_dim){
    x_init <- (3*l+i)*c_dim
  }else{
    x_init <- (((3*l+i)*c_dim)%%(c_dim*10))
  }
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    h <- l # 100 lines = 100 strata
    random_points <- rbind(random_points, list(a,b, h))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y", "Stratum")
return(random_points)
}
```

>>>>

---
title: "Testing bootstrap processes"
output: pdf_document
date: "2024-08-05"
---

@@

Idea for the bootstrap: for each stratum $h$, containing $n_h$ sampled points, $n_h - 1$ points are sampled with replacement. As a result, $100(n_h-1)$ points are sampled from the initial sample. In this case, 900 points are sampled.

@@

for (l in 0:((1000/c_dim)-1)) {
  if (l==0){
    x_init <- i*c_dim
  }else if ((3*l+i)*c_dim < (10)*c_dim){
    x_init <- (3*l+i)*c_dim
  }else{
    x_init <- (((3*l+i)*c_dim)%%(c_dim*10))
  }
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    h <- nrow(random_points)%%100 # 100 strata
    random_points <- rbind(random_points, list(a,b, h))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y", "Stratum")
return(random_points)
}
```

### 2024-09-12 Plots/Map_FR.Rmd
Better organized

```{r}
# Create a data frame with the coordinates and colors
# grid_data <- expand.grid(x = 1:30, y = 1:30)  # 10x10 grid
# grid_data$color <- "white"  # Default color for all squares

# (10--5)*2 = 30
# (51.5-41)*2 = 21

@@

  for (l in 0:20) {
    if (l==0){
      x_init <- i
    }else if ((3*l+i) < 10){
      x_init <- (3*l+i)
    }else{
      x_init <- (((3*l+i))%%(10))
    }
    k <- 0
    while (x_init + 10*k < 30) {
      a <- x_init + 10*k
      b <- l
      cells_x <- append(cells_x, a)
      # print(a)
      cells_y <- append(cells_y, b)
      # print(b)
      k <- k+1
    }
  }
  return(data.frame(x = cells_x, y = cells_y))
}

@@

# Highlight specific cells
highlighted_cells <- grid_sf[test$x + (test$y)*30, ]  # Example: highlight specific cells
highlighted_cells <- st_sf(geometry = highlighted_cells, color = "red")  # Assign color

# Combine highlighted and non-highlighted cells
grid_sf <- st_sf(geometry = grid_sf, color = "white")
grid_sf$color[which(grid_sf$geometry %in% highlighted_cells$geometry)] <- highlighted_cells$color
```

@@

# Plot the map
ggplot() +
  geom_sf(data = france, fill = "lightblue") +
  geom_sf(data = grid_sf, aes(fill = color), color = "black", alpha = 0.4) +
  scale_fill_identity() +  # Use the colors defined in the grid_sf object
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  theme_minimal() +
  ggtitle("Firs-stage sampling: sampling of cells within a grid") 
# +
  # annotation_scale(location = "bl") + # location = NULL
  # annotation_north_arrow(location = "bl")
```

@@

# Create a new column to determine land or water (or other features)
# Here, we'll simply mark if a grid cell intersects with land (France)
grid_sf$color2 <- ifelse(st_intersects(grid_sf, france, sparse = FALSE), "green", "blue")
```

@@

# Plot the map with color-coded grid cells
ggplot() +
  geom_sf(data = france, fill = "lightblue") +
  geom_sf(data = grid_sf, aes(fill = color3), color = "black", alpha = 0.4) +
  scale_fill_identity() +  # Use the colors defined in the grid_sf object
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  theme_minimal() 
# +
  # ggtitle("")
```

>>>>

```{r}
### Create a data frame with the coordinates and colors

# (10--5)*2 = 30 columns
# (51.5-41)*2 = 21 lines

@@

  for (l in 0:20) {
    if (l==0){
      x_init <- i
    }else if ((3*l+i) < 10){
      x_init <- (3*l+i)
    }else{
      x_init <- (((3*l+i))%%(10))
    }
    k <- 0
    while (x_init + 10*k < 30) {
      a <- x_init + 10*k
      b <- l
      cells_x <- append(cells_x, a)
      cells_y <- append(cells_y, b)
      k <- k+1
    }
  }
  return(data.frame(x = cells_x, y = cells_y))
}

@@

# Highlight specific cells
highlighted_cells <- grid_sf[test$x + (test$y)*30, ]  # highlight specific cells
highlighted_cells <- st_sf(geometry = highlighted_cells, color = "red")  # Assign color

# Combine highlighted and non-highlighted cells
grid_sf <- st_sf(geometry = grid_sf, color = "white") # Default color for all squares
grid_sf$color[which(grid_sf$geometry %in% highlighted_cells$geometry)] <- highlighted_cells$color
```

@@

# Plot the map
ggplot() +
  geom_sf(data = france, fill = "lightblue") +
  geom_sf(data = grid_sf, aes(fill = color), color = "black", alpha = 0.4) +
  scale_fill_identity() +  # Use the colors defined in the grid_sf object
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  theme_minimal() +
  ggtitle("Firs-stage sampling: sampling of cells within a grid")
```

@@

# Create a new column to determine land or water
# Here, we'll simply mark if a grid cell intersects with land (France)
grid_sf$color2 <- ifelse(st_intersects(grid_sf, france, sparse = FALSE), "green", "blue")
```

@@

# Plot the map with color-coded grid cells
ggplot() +
  geom_sf(data = france, fill = "lightblue") +
  geom_sf(data = grid_sf, aes(fill = color3), color = "black", alpha = 0.4) +
  scale_fill_identity() +  # Use the colors defined in the grid_sf object
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  theme_minimal()
```

### 2024-09-12 Results/All3_tab_algo6.Rmd
Better organized

In this document, \textbf{sampling through a grid} is used to create a sample and \textbf{three concentric circles} are used to compute local variables.

>>>>

In this document, \textbf{sampling through a grid} is used to create a sample, more precisely \textbf{algorithm 6}, and \textbf{three concentric circles} are used to compute local variables.

\textbf{Basic bootstrap} is used.
